### 2023-12-02 main.tex
lk

\noindent
\begin{remark}
\label{hole}
In the following we will often consider type $N$ as a distinguished type
(called `holes'), whose number at each site is fixed by the numbers of particles 
of the remaining $N-1$ types. 
% It is interesting to observe that, in case we stop distinguishing the types of the ``true'' particles of types 
% (those of types $\{1,\ldots,N-1\}$), we retrieve the 
% boundary-driven version of the partial symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
In the case $N=2$, we retrieve the  boundary-driven version of the symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
% \todo{I would prefer to say this is for N=2}
The choice of type $N$ for the distinguished type is arbitrary and does not affect
the results. However, we shall see in Section~\ref{sectionDuality}, that
the dual process breaks the symmetry.
In particular, in the  long-time limit the dual process has only particles of type $N$ on the graph.
Interpreting type $N$ as an hole, one can then say that the dual process voids the graph.  
\todo{Exchange hole with void? Also can the remark be just text?}
\end{remark}

@@

If condition \eqref{reversibilityCondition} is not met, then in general reversibility is lost (but see Remark~\ref{rem-rev} below): 

@@

		\todo{?}

@@

\subsection{Integrable process on a line segment}\label{subsection-description-process-LINE}
In this section we specialize the multispecies stirring process to the geometry of the one-dimensional chain with sites $\{1,\ldots,L\}$ where two reservoirs act at the boundary sites $1$ and $L$.  The reservoirs, exchanging particle with the external environment, put the chain out of equilibrium. This process was studied previously in \cite{vanicat2017exact}.

The Hamiltonian that we consider here is obtained from \eqref{OriginalHamiltonian} assuming that the conductances are
\begin{equation}
	\omega_{x,y}=\begin{cases}
		1 \quad \text{if}\quad |x-y|=1\\
		0\quad \text{otherwise}
	\end{cases}
\end{equation}
and the coupling to reservoirs are
\begin{equation}
	\Gamma_{x}=\begin{cases}
		1\quad \text{if} \quad x\in \{1,L\}\\
		0\quad \text{otherwise}
	\end{cases}\dt
\end{equation}
The case $\nu=1$, i.e. one particle at most for each site, is special. We will consider this extensively in Section~\ref{sectionIntegrabiliy}.
We will denote the element of the representation of the ${gl}(N)$  Lie algebra with $\nu=1$ by lowercase letters, i.e. $(e_{ab})_{a,b\in\{1,\ldots,N\}}$, to highlight that we are in the first fundamental representation of the algebra.
The Hamiltonian now reads
\begin{equation}\label{hamiltonian}
	H=H_{\text{left}}+H_{\text{bulk}}+H_{\text{right}}
\end{equation}
where
\begin{equation}
	H_{\text{bulk}}=\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}\dt
\end{equation}
Here $\mathcal{H}_{x,x+1}$  is a copy (acting on the vertices of the edge $x,x+1$) of the two-site Hamiltonian
\begin{equation}\label{H-corsivo}
	\begin{split}
		\mathcal{H}=P-\id
	\end{split}
\end{equation}
with 
\begin{equation}
	P=\sum_{a,b=1}^Ne_{ab}\otimes e_{ba}.
\end{equation} 
 In this context it will be useful to introduce the following notation for the occupation variable of the process. Each configuration is denoted by
\begin{equation}\label{Tau-Notation}
	|\bm{\tau}\rangle=|\tau_{1},\ldots,\tau_{L}\rangle
\end{equation} 
with $\tau_{x}\in \{1,\ldots,N\}$, $\forall x\in \{1,\ldots,L\}$. Since the maximal occupancy at each site is $\nu=1$,  the configuration $\bm{n}$ introduced in Section~\ref{subsectionGeneratorStr} and $\bm{\tau}$ are related by 
\begin{equation}\label{notation-change-relation}
	n_{a}^{x}=\delta_{\tau_{x},a}\dt
\end{equation}
The state space of the process is now 
\begin{equation}
	\Omega^{'}=\left\{(\tau_{1},\ldots,\tau_{L})\;:\; \tau_{x}\in\{1,\ldots,N\}\right\}\dt
\end{equation}
The action of $\mathcal{H}$ amounts to the permutation of the occupations of the two sites, i.e.
\begin{equation}
	\begin{split}
		\mathcal{H}| \tau\rangle\otimes   |\tau'\rangle&=|\tau'\rangle \otimes |\tau\rangle-|\tau\rangle \otimes|\tau'\rangle\dt
	\end{split}
\end{equation}
The boundary terms are given by 
\begin{equation}
	H_{\text{left}}=\begin{pmatrix}
		\alpha_{1}-1&\alpha_{1}&\alpha_{1}&\ldots&\ldots&\alpha_{1}\\
		\alpha_{2}&\alpha_{2}-1&\alpha_{2}&\ldots&\ldots&\alpha_{2}\\
		\vdots&\vdots& &\ddots& &\vdots\\
		\alpha_{N-1}&\alpha_{N-1}&\ldots&\ldots&\alpha_{N-1}-1&\alpha_{N-1}\\
		\alpha_{N}&\alpha_{N}&\ldots&\ldots&\alpha_{N}&\alpha_{N}-1
	\end{pmatrix}
\end{equation}
and 
\begin{equation}
	H_{\text{right}}=\begin{pmatrix}
		\beta_{1}-1&\beta_{1}&\beta_{1}&\ldots&\ldots&\beta_{1}\\
		\beta_{2}&\beta_{2}-1&\beta_{2}&\ldots&\ldots&\beta_{2}\\
		\vdots&\vdots& &\ddots& &\vdots\\
		\beta_{N-1}&\beta_{N-1}&\ldots&\ldots&\beta_{N-1}-1&\beta_{N-1}\\
		\beta_{N}&\beta_{N}&\ldots&\ldots&\beta_{N}&\beta_{N}-1
	\end{pmatrix}
\end{equation}
where, without loss of generality,  we assume that the parameters satisfy
\begin{equation}\label{ratesConditions}
	\sum_{a=1}^{N}\alpha_{a}=1,\qquad\sum_{a=1}^{N}\beta_{a}=1\;.
\end{equation} 
%We define also 
%\begin{equation}\label{lambdaConditions}
%	(\alpha_{a}-\beta_{a})=\alpha_{a}-\beta_{a}\quad\text{with}\quad \sum_{a=1}^{N}(\alpha_{a}-\beta_{a})=0\,.
%\end{equation}
Under these assumption the Hamiltonian~\ref{hamiltonian} is integrable (see \cite{vanicat2017exact} for its derivation within the Quantum Inverse Scattering method).

>>>>

\noindent
\begin{remark}
\label{hole}
In the following we will often consider type $N$ as a distinguished type
(called `holes'), whose number at each site is fixed by the numbers of particles 
of the remaining $N-1$ types. 
% It is interesting to observe that, in case we stop distinguishing the types of the ``true'' particles of types 
% (those of types $\{1,\ldots,N-1\}$), we retrieve the 
% boundary-driven version of the partial symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
In the case $N=2$, we retrieve the  boundary-driven version of the symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
% \todo{I would prefer to say this is for N=2}
The choice of type $N$ for the distinguished type is arbitrary and does not affect
the results. However, we shall see in Section~\ref{sectionDuality}, that
the dual process breaks the symmetry.
In particular, in the  long-time limit the dual process has only particles of type $N$ on the graph.
Interpreting type $N$ as an hole, one can then say that the dual process voids the graph.   
\end{remark}

@@

If condition \eqref{reversibilityCondition} is not met, then in general reversibility is lost: 

@@

% 		\todo{?}

@@

\subsection{Integrable process on a line segment}\label{subsection-description-process-LINE}
In this section we specialize the multispecies stirring process to the geometry of the one-dimensional chain with sites $\{1,\ldots,L\}$ where two reservoirs are attached to the boundary sites $1$ and $L$.  The reservoirs, exchanging particles with the external environment, put the chain out of equilibrium.  The  Hamiltonian that we consider here is obtained from \eqref{OriginalHamiltonian} assuming that the conductances are
\begin{equation}
	\omega_{x,y}=\begin{cases}
		1 \quad \text{if}\quad |x-y|=1\\
		0\quad \text{otherwise}
	\end{cases}
\end{equation}
and the coupling to reservoirs are
\begin{equation}
	\Gamma_{x}=\begin{cases}
		1\quad \text{if} \quad x\in \{1,L\}\\
		0\quad \text{otherwise}
	\end{cases}\dt
\end{equation}
The case $\nu=1$, i.e. one particle at most for each site, is integrable and has been considered previously in \cite{vanicat2017exact}. 
For  $\nu=1$ we denote the ${gl}(N)$  generators   as  $(e_{ab})_{a,b\in\{1,\ldots,N\}}$.
The Hamiltonian can then be written as
\begin{equation}\label{hamiltonian}
	H=H_{\text{left}}+H_{\text{bulk}}+H_{\text{right}}
\end{equation}
where
\begin{equation}
	H_{\text{bulk}}=\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}\dt
\end{equation}
Here $\mathcal{H}_{x,x+1}$  is a copy (acting on the vertices of the edge $x,x+1$) of the two-site Hamiltonian
\begin{equation}\label{H-corsivo}
	\begin{split}
		\mathcal{H}=P-\id
	\end{split}
\end{equation}
with the permutation matrix
\begin{equation}
	P=\sum_{a,b=1}^Ne_{ab}\otimes e_{ba}.
\end{equation} 
 In this context it will be useful to introduce the following notation for the occupation variables of the process. Each configuration is denoted by
\begin{equation}\label{Tau-Notation}
	|\bm{\tau}\rangle=|\tau_{1},\ldots,\tau_{L}\rangle
\end{equation} 
with $\tau_{x}\in \{1,\ldots,N\}$, $\forall x\in \{1,\ldots,L\}$. Since the maximal occupancy at each site is $\nu=1$,  the configuration $\bm{n}$ introduced in Section~\ref{subsectionGeneratorStr} and $\bm{\tau}$ are related by 
\begin{equation}\label{notation-change-relation}
	n_{a}^{x}=\delta_{\tau_{x},a}\dt
\end{equation}
The state space of the process is now 
\begin{equation}
	\Omega^{'}=\left\{(\tau_{1},\ldots,\tau_{L})\;:\; \tau_{x}\in\{1,\ldots,N\}\right\}\dt
\end{equation}
The action of $\mathcal{H}$ amounts to the permutation of the occupations of the two sites, i.e.
\begin{equation}
	\begin{split}
		\mathcal{H}| \tau\rangle\otimes   |\tau'\rangle&=|\tau'\rangle \otimes |\tau\rangle-|\tau\rangle \otimes|\tau'\rangle\dt
	\end{split}
\end{equation}
The boundary terms are given by 
\begin{equation}
	H_{\text{left}}=\begin{pmatrix}
		\alpha_{1}-1&\alpha_{1}&\alpha_{1}&\ldots&\ldots&\alpha_{1}\\
		\alpha_{2}&\alpha_{2}-1&\alpha_{2}&\ldots&\ldots&\alpha_{2}\\
		\vdots&\vdots& &\ddots& &\vdots\\
		\alpha_{N-1}&\alpha_{N-1}&\ldots&\ldots&\alpha_{N-1}-1&\alpha_{N-1}\\
		\alpha_{N}&\alpha_{N}&\ldots&\ldots&\alpha_{N}&\alpha_{N}-1
	\end{pmatrix}
\end{equation}
and 
\begin{equation}
	H_{\text{right}}=\begin{pmatrix}
		\beta_{1}-1&\beta_{1}&\beta_{1}&\ldots&\ldots&\beta_{1}\\
		\beta_{2}&\beta_{2}-1&\beta_{2}&\ldots&\ldots&\beta_{2}\\
		\vdots&\vdots& &\ddots& &\vdots\\
		\beta_{N-1}&\beta_{N-1}&\ldots&\ldots&\beta_{N-1}-1&\beta_{N-1}\\
		\beta_{N}&\beta_{N}&\ldots&\ldots&\beta_{N}&\beta_{N}-1
	\end{pmatrix}
\end{equation}
where, without loss of generality,  we assume that the parameters satisfy
\begin{equation}\label{ratesConditions}
	\sum_{a=1}^{N}\alpha_{a}=1,\qquad\sum_{a=1}^{N}\beta_{a}=1\;.
\end{equation} 
%We define also 
%\begin{equation}\label{lambdaConditions}
%	(\alpha_{a}-\beta_{a})=\alpha_{a}-\beta_{a}\quad\text{with}\quad \sum_{a=1}^{N}(\alpha_{a}-\beta_{a})=0\,.
%\end{equation}
Under these assumption the Hamiltonian~\ref{hamiltonian} is integrable (see \cite{vanicat2017exact} for its derivation within the Quantum Inverse Scattering method).

We will consider it extensively in Section~\ref{sectionIntegrabiliy}.
